What We Eval in the Shadows

### 2021-04-05 paper/main.tex
Fix
Removed some unused text

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations
\usepackage{listings,hyperref,multirow,paralist,xspace,url,wrapfig}
\usepackage{tabularx}
\input{macro}
\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}
\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}

@@

@@

This section focuses on CRAN packages and reports statistics about the usage of
\eval. There are \PkgEvalCallSites call sites to \eval in \PkgPackages
packages.  These represent \PkgPackagesRatio of all packages or \PkgCodeRatio
of the entire code size of CRAN. Over half of the packages have less than 3
\eval call sites and except \MaxEvalCallSitesPackage (\MaxEvalCallSitesCount
\eval calls), all packages contain less than \MaxEvalCallSitesRest \eval call
sites (\cf Fig~\ref{fig:pkgs-eval-callsites-hist}) These \eval calls appear in
\PkgFunsWithEval R functions (\CranFunsWithEvalRatio of all R functions in
CRAN).

@@

\begin{figure}[t]
\begin{minipage}{.3\textwidth}
  \vspace*{14mm}
  \hspace*{-10mm}
  \includegraphics[width=1.5\textwidth]{pkgs-eval-callsites-hist.pdf} \caption{CRAN
  \eval call sites}%
  \label{fig:pkgs-eval-callsites-hist}
\end{minipage}
\begin{minipage}{.6\textwidth}
  \centering
  \includegraphics[width=\textwidth]{traced-eval-callsites.pdf}
  \caption{\eval call sites coverage of the \PkgPackages packages.}%
  \label{fig:traced-eval-callsites}
\end{minipage}
\end{figure}

@@

To estimate how much executable code is injected through \eval, we measure the
the number of abstract syntax tree nodes as returned by the parser; for example,
\c{x+1} counts as 3.\footnote{We attempted to measure string lengths after
de-parsing expressions, but these measurements were dominated by de-parsed data
objects, which could range in the MBs.} The median argument size is
\Medianszeval (due to the many symbols) and the average is \Avgszeval nodes. The
largest \eval input observed is \Maxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes.
%
\begin{wrapfigure}{r}{5.8cm}
 \hspace*{-10mm}
 \includegraphics[width=0.5\textwidth, trim=4.8cm 0 6cm 0, clip]{size_loaded_distribution}
 \caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}
%
The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
from the graph for legibility.

@@

\paragraph{Base}
We recorded \baseAllcalls \eval calls in the base libraries with fewer than 4\%
of those being to any variants. Unlike CRAN, Base calls \c{evalq} (\baseEvalqs).
Out of \baseStaticeval call sites, \baseTriggeredevalpct are exercised. Most
arguments, \baseCodepercent, are expressions. However, unlike CRAN,
\baseLanguagepercent of the arguments are language objects. The median argument
size is \baseMedianszeval, which is more than for CRAN. The maximum size is
\baseMaxszeval. Small instruction counts, less than 50, amount for
\baseSmalleventspct of calls.

\paragraph{Kaggle}
Only \kaggleNbruns scripts trigger \eval. In total, \kaggleAllcalls \eval were
recorded; none are calls to variants. Out of \kaggleStaticeval call sites,
\kaggleTriggeredeval were exercised. Most call sites are invoked only once. Only
one site is called \kaggleMaxcalls times. Expressions amount for
\kaggleCodepercent of arguments. Unlike CRAN corpus, there are very few symbols
(\kaggleSymbolpercent). Most expressions result from calls to \c{parse}, thus
most \evals start with strings. The median argument size is \kaggleMedianszeval,
Which makes sense, as few arguments are only symbols. The largest argument is
\kaggleMaxszeval. The distribution of instructions per eval is similar to CRAN;
\kaggleSmalleventspct of \evals execute fewer than 50 instructions.

@@

Most of the time, \ie in \packageEnvirEnvSitePercent of the sites, \eval is
passed an actual environment.

In \packageNbEnvirIncorrectSites sites (\packageEnvirIncorrectSitePercent),
\eval is passed an incorrect argument, for instance \c{NA}, a boolean, which
makes \eval fail.

@@

\paragraph{Provenance.} 
% Explain tha some generate expression objects and other language objects?
There are various ways to obtain expressions that \eval will evaluate:
\begin{itemize}
    \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
    \item \c{match.call} captures a call and its arguments as expressions.
    \item \c{quote}, \c{enquote} returns its argument unevaluated, as an R expression.
    \item \c{expression}  creates a vector of expression
    objects from text.
    \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
    \item \c{comp} compiles R code into R bytecode.
\end{itemize}

@@

\mypara{New Bindings.} Find number of times, eval introduces new bindings. There
are many ways -- library, load, attach, source, and explicitly introduce
bindings using super-assign, assign and define.

\mypara{C Calls.} Find number of times, eval makes call to native code using -
.Call, .External, etc. Find out if there are native functions that are called
only from within eval.

\mypara{Non-local returns} Eval can do non-local returns effective bypassing
evaluation of the rest of the function. For instance, it can inject a \c{break} or \c{next} instructions in a loop. It is very rare however, as it happens only twice in our dataset. % What kind of non-local returns?
% eval(quote(return(89))) does not work

\mypara{Purity.} Conclude with number of evals that are ``pure'', i.e. evals
which could be ignored by a static analyzer without any problem.

@@

We present here the main uses of \eval we have witnessed in R. Sometimes, they are design patterns, sometimes, \eval could be replaced by more specific functions. For each use of \eval, we discuss whether it is necessary to use.

\subsection{Variable lookup}

\eval is used to look-up the value of a variable in a given environment, different from the current scope so the \c{envir} argument of \eval is explicitly specified.

% TODO: numbers on the prevalence of the pattern

\eval is not necessary in that case and can be replaced by \c{get}, as follows:

\begin{lstlisting}
 eval(quote(x), envir = env)
\end{lstlisting}
$\Downarrow$
\begin{lstlisting}
get("x", envir = env)
\end{lstlisting}

% This is actually linked to NSE, as the symbol given to \eval usually comes
% from match.call or susbtitute

\subsection{Non-Standard Evaluation} % Or metaprogramming

\subsection{\c{substitute}}

\subsection{\c{match.call}}

\subsection{Logging}

\subsection{Code transform}

% Derivation of symbolic expressions

\subsection{Plotting}

\section{\eval expressivity and macros}

@@

\section{The dangers of \eval} \label{sec:dangerseval}

\subsection{Eval Usage in R}

>>>>

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations
\usepackage{listings,hyperref,multirow,paralist,xspace,url,wrapfig}
\usepackage{tabularx}
\input{macro}
\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}
\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}
\newcommand{\authorcomment}[3]{\xspace\textcolor{#1}{{\bf #2} #3}\xspace}
\newcommand{\todo}[1]{\authorcomment{red}{TODO}{#1}}

@@

\todo{Look at Improving dynamic code analysis by code abstraction, VPT'21}

@@

\begin{wrapfigure}{r}{5cm} \hspace*{-12mm}
  \centering
  \includegraphics[width=75mm]{pkgs-eval-callsites-hist.pdf} \caption{CRAN
  \eval call sites}%
  \label{fig:pkgs-eval-callsites-hist}
\end{wrapfigure}
%
This section focuses on CRAN packages and reports statistics about the usage of
\eval. There are \PkgEvalCallSites call sites to \eval in \PkgPackages
packages.  These represent \PkgPackagesRatio of all packages or \PkgCodeRatio
of the entire code size of CRAN. Over half of the packages have less than 3
\eval call sites and except \MaxEvalCallSitesPackage (\MaxEvalCallSitesCount
\eval calls), all packages contain less than \MaxEvalCallSitesRest \eval call
sites (\cf Fig~\ref{fig:pkgs-eval-callsites-hist}) These \eval calls appear in
\PkgFunsWithEval R functions (\CranFunsWithEvalRatio of all R functions in
CRAN).

@@

\begin{figure}[tb]
  \centering
  \includegraphics[width=.8\textwidth]{traced-eval-callsites.pdf}
  \centering
  \caption{\eval call sites coverage of the \PkgPackages packages.}%
  \label{fig:traced-eval-callsites}
\end{figure}

@@

To estimate how much executable code is injected through \eval, we measure the
the number of abstract syntax tree nodes as returned by the parser; for example,
\c{x+1} counts as 3.\footnote{We attempted to measure string lengths after
de-parsing expressions, but these measurements were dominated by de-parsed data
objects, which could range in the MBs.} The median argument size is
\Medianszeval (due to the many symbols) and the average is \Avgszeval nodes. The
largest \eval input observed is \Maxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes.
The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
from the graph for legibility.
%
\begin{wrapfigure}{r}{5.8cm}
 \hspace*{-10mm}
 \includegraphics[width=0.5\textwidth, trim=4.8cm 0 6cm 0, clip]{size_loaded_distribution}
 \caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}
%

@@

\paragraph{Base}
We recorded \baseAllcalls \eval calls in the base libraries with fewer than 4\%
of those being to any variants. Unlike CRAN, Base calls \c{evalq} (\baseEvalqs).
Out of \baseStaticeval call sites, \baseTriggeredevalpct are exercised. Most
arguments, \baseCodepercent, are expressions. However, unlike CRAN,
\baseLanguagepercent of the arguments are language objects. The median argument
size is \baseMedianszeval, which is more than for CRAN. The maximum size is
\baseMaxszeval. Small instruction counts, less than 50, amount for
\baseSmalleventspct of calls.
\todo{More info about base and conclude it here.}

\paragraph{Kaggle}
Only \kaggleNbruns scripts trigger \eval. In total, \kaggleAllcalls \eval were
recorded; none are calls to variants. Out of \kaggleStaticeval call sites,
\kaggleTriggeredeval were exercised. Most call sites are invoked only once. Only
one site is called \kaggleMaxcalls times. Expressions amount for
\kaggleCodepercent of arguments. Unlike CRAN corpus, there are very few symbols
(\kaggleSymbolpercent). Most expressions result from calls to \c{parse}, thus
most \evals start with strings. The median argument size is \kaggleMedianszeval,
Which makes sense, as few arguments are only symbols. The largest argument is
\kaggleMaxszeval. The distribution of instructions per eval is similar to CRAN;
\kaggleSmalleventspct of \evals execute fewer than 50 instructions.
\todo{More info about kaggle and conclude it here.}

@@

\todo{Pierre: check the numbers - we want to know when it is an environment, when it is a list / data frame}
Most of the time, \ie in \packageEnvirEnvSitePercent of the sites, \eval is
passed an actual environment.

@@

\subsection{Provenance.}
% Explain tha some generate expression objects and other language objects?
There are various ways to obtain expressions that \eval will evaluate:
\begin{itemize}
    \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
    \item \c{match.call} captures a call and its arguments as expressions.
    \item \c{quote}, \c{enquote} returns its argument unevaluated, as an R expression.
    \item \c{expression}  creates a vector of expression
    objects from text.
    \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
    \item \c{comp} compiles R code into R bytecode.
\end{itemize}

@@

\subsection{Side effects in \eval}

\todo{Filip: discuss the side-effects}

@@

\todo{Illustrate things that we have discussed in Section 5 - examples}
\todo{Things to discuss: model.frame, match.call, substitute, parse, plots}

% We present here the main uses of \eval we have witnessed in R. Sometimes, they are design patterns, sometimes, \eval could be replaced by more specific functions. For each use of \eval, we discuss whether it is necessary to use.

% \subsection{Variable lookup}

% \eval is used to look-up the value of a variable in a given environment, different from the current scope so the \c{envir} argument of \eval is explicitly specified.

% \subsection{Metaprogramming}

% \mypara{\c{substitute}}

% \mypara{\c{match.call}}

% \todo{Talk about model.frame}

% \subsection{Logging}

% \subsection{Code transform}

% % Derivation of symbolic expressions

% \subsection{Plotting}

% \section{\eval expressivity and macros}

@@
